---
title: "Agent"
format:
  html:
    toc: true
execute:
  echo: true
  warning: false
  error: false
---

```{r}
#| include: false
if (requireNamespace("pkgload", quietly = TRUE)) pkgload::load_all()
library(mini007)

# helper for examples; adjust your API key retrieval
retrieve_open_ai_credential <- function() {
  Sys.getenv("OPENAI_API_KEY")
}

openai_4_1_mini <- ellmer::chat(
  name = "openai/gpt-4.1-mini",
  credentials = retrieve_open_ai_credential,
  echo = "none"
)
```

# Agent

## Creating an Agent

An Agent is built upon an LLM object created by the `ellmer` package, in the
following examples, we'll work with the `OpenAI` models, however you can use any
model/combination of models you want:

```{r}
polar_bear_researcher <- Agent$new(
  name = "POLAR BEAR RESEARCHER",
  instruction = "You are an expert in polar bears, you task is to collect information about polar bears. Answer in 1 sentence max.",
  llm_object = openai_4_1_mini
)
```

Each created Agent has an `agent_id` (among other meta information):

```{r}
polar_bear_researcher$agent_id
```

At any time, you can tweak the `llm_object`:

```{r}
polar_bear_researcher$llm_object
```

An agent can provide the answer to a prompt using the `invoke` method:

```{r}
# Note: the following call will actually hit the LLM if the key is set
# and may incur cost.
polar_bear_researcher$invoke("Are polar bears dangerous for humans?")
```

You can also retrieve a list that displays the history of the agent:

```{r}
polar_bear_researcher$messages
```
Or the `ellmer` way:

```{r}
polar_bear_researcher$llm_object
```

## Agents and Messages

### Managing Agent Conversation History

The `clear_and_summarise_messages` method allows you to compress an agent's
conversation history into a concise summary and clear the message history while
preserving context. This is useful for maintaining memory efficiency while
keeping important conversation context.

```{r}
polar_bear_researcher$clear_and_summarise_messages()
polar_bear_researcher$messages
```

This method summarises all previous conversations into a paragraph and appends
it to the system prompt, then clears the conversation history. The agent
retains the context but with reduced memory usage.

### Keep only the most recent messages

When a conversation grows long, you can keep just the last N messages while
preserving the system prompt. This helps control token usage without fully
resetting context.

```{r}
openai_4_1_mini <- ellmer::chat(
  name = "openai/gpt-4.1-mini",
  credentials = retrieve_open_ai_credential,
  echo = "none"
)

agent <- Agent$new(
  name = "history_manager",
  instruction = "You are a concise assistant.",
  llm_object = openai_4_1_mini
)

agent$invoke("What is the capital of Italy?")
agent$invoke("What is the capital of Germany?")
agent$invoke("What is the capital of Algeria?")
agent$messages
```

```{r}
# Keep only the last 2 messages (system prompt is preserved)
agent$keep_last_n_messages(n = 2)
agent$messages
```

### Manually Adding Messages to an Agent’s History

You can inject any message (system, user, or assistant) directly into an
Agent’s history with `add_message(role, content)`. This is helpful to
reconstruct, supplement, or simulate conversation steps.

```{r}
openai_4_1_mini <- ellmer::chat(
  name = "openai/gpt-4.1-mini",
  credentials = retrieve_open_ai_credential,
  echo = "none"
)
agent <- Agent$new(
  name = "Pizza expert",
  instruction = "You are a Pizza expert",
  llm_object = openai_4_1_mini
)

# Add a user message, an assistant reply, and a system instruction:
agent$add_message("user", "Where can I find the best pizza in the world?")
agent$add_message("assistant", "You can find the best pizza in the world in Algiers, Algeria. It's tasty and crunchy.")

# View conversation history
agent$messages
```

```{r}
agent$invoke("summarise the previous conversation")
```

A subset of the examples from the README.  Additional functionality such as
`generate_and_register_tool()` and planning is covered in the LeadAgent page.

