---
title: "mini007"
format:
  html:
    toc: true
execute:
  echo: true
  warning: false
  error: false
---

```{r}
#| include: false
# load the development version of the package so examples run locally
if (requireNamespace("pkgload", quietly = TRUE)) pkgload::load_all()
library(mini007)
```

# Welcome to **mini007**

`mini007` provides a lightweight and extensible framework for multi-agents orchestration
processes capable of decomposing complex tasks and assigning them to specialized agents.

> Each `agent` is an extension of an `ellmer` object. `mini007` relies heavily on the
> excellent `ellmer` package but aims to make it easy to create a process where multiple
> specialized agents help each other sequentially in order to execute a task.

## Highlights

- 🧠 Memory and identity for each agent via `uuid` and message history.
- ⚙️ Built-in task decomposition and delegation via `LLM`.
- 🔄 Agent-to-agent orchestration with result chaining.
- 🌐 Compatible with any chat model supported by `ellmer`.
- 🧑 Possibility to set a Human In The Loop (`HITL`) at various execution steps

## Installation

You can install `mini007` from CRAN with:

```{r}
#| eval: false
install.packages("mini007")
```

Or, during development, load the package as shown in the setup chunk above.


## Get started

Navigate to the **Agents** page to see how to create and interact with individual
agents, or jump straight to the **LeadAgent** page for multi‑agent orchestration.


