---
title: "mini007"
format:
  html:
    toc: true
execute:
  echo: true
  warning: false
  error: false
---

```{r}
#| include: false
if (requireNamespace("pkgload", quietly = TRUE)) pkgload::load_all()
library(mini007)
```


# Welcome to **mini007**

`mini007` is a lightweight and extensible R framework for building multi-agent
AI systems. It provides three complementary components — `Agent`,
`LeadAgent`, and `Workflow` — that can be used independently or combined to
tackle complex tasks that benefit from specialised, collaborating LLM workers.

> `mini007` relies heavily on the excellent
> [`ellmer`](https://ellmer.tidyverse.org/) package and is compatible with any
> chat model it supports: OpenAI, Anthropic, Google, Ollama, and more.

## Highlights

- 🧠 **Persistent memory**: every agent maintains its full conversation
  history with export, import, and summarisation support.
- ⚙️ **Task decomposition**: `LeadAgent` automatically splits a complex
  prompt into subtasks and delegates each one to the most suitable agent.
- 🔀 **Sequential pipelines**: `Workflow` gives you explicit control over
  execution order, branching logic, and which handler runs at each step.
- 💾 **Caching**: workflow station results are cached by input so subsequent
  runs skip redundant LLM calls.
- 🔧 **Tool support**: register `ellmer` tools on any agent, or generate new
  ones from a natural-language description.
- 💰 **Budget control**: set a USD spending cap on any agent with a
  configurable policy (`abort`, `warn`, or `ask`).
- 🧑 **Human In The Loop**: pause at specific steps so a
  human can review, edit, or stop the response before it continues.
- 🌐 **Model-agnostic**: swap the underlying LLM without changing any
  orchestration logic.

## Installation

```{r}
#| eval: false
install.packages("mini007")
```

## Key components

### `Agent`

An `Agent` wraps an `ellmer` chat object and adds identity (UUID), persistent
message history, budget tracking, and tool support. It is the fundamental
building block of the framework.

```{r}
#| eval: false
retrieve_open_ai_credential <- function() {
  Sys.getenv("OPENAI_API_KEY")
}

llm <- ellmer::chat(
  name        = "openai/gpt-4.1-mini",
  credentials = retrieve_open_ai_credential,
  echo        = "none"
)
```

```{r}
#| eval: false
researcher <- Agent$new(
  name        = "researcher",
  instruction = "Answer factual questions concisely and accurately.",
  llm_object  = llm
)

researcher$invoke("What is the boiling point of water at high altitude?")
```

Agents remember the conversation, so follow-up questions work naturally:

```{r}
#| eval: false
researcher$invoke("Why does it change with altitude?")
```

Notable methods:

| Method | Purpose |
|---|---|
| `$invoke(prompt)` | Send a prompt and get a response |
| `$keep_last_n_messages(n)` | Trim history to the last *n* turns |
| `$share_context_with(agent)` | Copy recent messages to another agent |
| `$set_budget(amount)` | Set a USD spending limit |
| `$register_tools(...)` | Attach `ellmer` tool objects |
| `$generate_and_register_tool(desc)` | Generate a tool from plain English |
| `$generate_execute_r_code()` | Write, validate, and run R code |
| `$validate_response(response, criteria)` | LLM-based output verification |
| `$clone_agent()` | Create an independent copy with a new UUID |
| `$export_messages_history(path)` | Save conversation to JSON |
| `$load_messages_history(path)` | Restore conversation from JSON |

---

### `LeadAgent`

A `LeadAgent` extends `Agent` with multi-agent orchestration: it decomposes a
complex prompt into subtasks and routes each one to the most suitable
registered agent, collecting and chaining the results.

```{r}
#| eval: false
summariser <- Agent$new(
  name        = "summariser",
  instruction = "Summarise the text you receive into three bullet points.",
  llm_object  = llm
)

translator <- Agent$new(
  name        = "translator",
  instruction = "Translate the text you receive from English to French.",
  llm_object  = llm
)

lead <- LeadAgent$new(name = "Lead", llm_object = llm)
lead$register_agents(c(summariser, translator))

lead$invoke("Summarise the history of the Sahara Desert, then translate it into French.")
```

Notable methods:

| Method | Purpose |
|---|---|
| `$register_agents(agents)` | Add agents to the pool |
| `$generate_plan(prompt)` | Preview task-to-agent assignments before running |
| `$visualize_plan()` | Render the plan as an interactive directed graph |
| `$invoke(prompt)` | Decompose, delegate, and return the final answer |
| `$broadcast(prompt)` | Send one prompt to all agents and collect every response |
| `$judge_and_choose_best_response(prompt)` | Let the lead pick the best answer from all agents |
| `$agents_dialog(prompt, id1, id2)` | Run an iterative negotiation between two agents |
| `$set_hitl(steps)` | Pause at specified steps for human review or editing |

---

### `Workflow`

A `Workflow` is a predefined, sequential pipeline of **Stations** connected by
**Routes**. Each Station's output feeds directly into the next one's input. You
decide the order, the branching conditions, and what runs at each step — the
handler can be an `Agent`, a `WorkflowAgent`, or a plain R function.

```{r}
#| eval: false
# Plain R function as a station — no LLM needed
clean <- function(text) trimws(gsub("\\s+", " ", text))

wf <- Workflow$new(name = "article-pipeline", use_cache = TRUE)

wf$add_station("clean",    clean)
wf$add_station("write",    Agent$new(
  name        = "writer",
  instruction = "Turn the notes you receive into an engaging paragraph.",
  llm_object  = llm
))
wf$add_station("edit",     Agent$new(
  name        = "editor",
  instruction = "Polish the paragraph for grammar and conciseness.",
  llm_object  = llm
))

wf$add_route("clean", "write")
wf$add_route("write", "edit")

wf$run("  The  history  of   the  printing   press  ")
```

Routes can carry a condition function to create branching pipelines:

```{r}
#| eval: false
wf$add_route(
  from      = "classify",
  to        = "positive_reply",
  condition = function(out) grepl("positive", tolower(out))
)
```

A finished `Workflow` can be wrapped as a `WorkflowAgent` and registered with
a `LeadAgent` or nested inside another `Workflow`:

```{r}
#| eval: false
wa <- wf$as_agent(name = "article_agent")
lead$register_agents(list(wa))
```

Notable methods:

| Method | Purpose |
|---|---|
| `$add_station(name, handler)` | Register a Station (agent or function) |
| `$add_route(from, to, condition)` | Connect two Stations, optionally gated by a condition |
| `$set_entry(name)` | Set the starting Station |
| `$run(input)` | Execute the pipeline sequentially |
| `$clear_cache()` | Invalidate cached Station results |
| `$set_hitl(steps)` | Pause at specified steps for human review or editing |
| `$as_agent(name, instruction)` | Wrap the workflow as a `WorkflowAgent` |
| `$visualize()` | Render the pipeline as an interactive directed graph |

## Get started

Navigate through the pages in the top navigation bar:

- **Agents** — individual agent creation, memory, tools, budgets, and code generation.
- **LeadAgent** — multi-agent orchestration, broadcasting, HITL, and agent dialog.
- **Tools** — building and registering custom tools.
- **Workflow** — sequential pipelines, conditional routing, caching, and nesting.
- **Testing** — utilities for writing tests against LLM responses.
