---
title: "LeadAgent"
format:
  html:
    toc: true
execute:
  echo: true
  warning: false
  error: false
---

```{r}
library(mini007)

retrieve_open_ai_credential <- function() {
  Sys.getenv("OPENAI_API_KEY")
}

openai_4_1_mini <- ellmer::chat(
  name = "openai/gpt-4.1-mini",
  credentials = retrieve_open_ai_credential,
  echo = "none"
)
```

## Creating a multi-agents orchestraction 

We can create as many Agents as we want, the `LeadAgent` will dispatch the instructions to the agents and provide with the final answer back. Let's create three Agents, a `researcher`, a `summarizer` and a `translator`: 


```{r}

researcher <- Agent$new(
  name = "researcher",
  instruction = "You are a research assistant. Your job is to answer factual questions with detailed and accurate information. Do not answer with more than 2 lines",
  llm_object = openai_4_1_mini
)

summarizer <- Agent$new(
  name = "summarizer",
  instruction = "You are agent designed to summarise a give text into 3 distinct bullet points.",
  llm_object = openai_4_1_mini
)

translator <- Agent$new(
  name = "translator",
  instruction = "Your role is to translate a text from English to German",
  llm_object = openai_4_1_mini
)
```


Now, the most important part is to create a `LeadAgent`: 

```{r}
lead_agent <- LeadAgent$new(
  name = "Leader", 
  llm_object = openai_4_1_mini
)
```

Note that the `LeadAgent` cannot receive an `instruction` as it has already the necessary instructions. 

Next, we need to assign the Agents to `LeadAgent`, we do it as follows: 

```{r}
lead_agent$register_agents(c(researcher, summarizer, translator))

lapply(lead_agent$agents, function(x) {x$name})
```

Before executing your prompt, you can ask the `LeadAgent` to generate a plan so that you can see which `Agent` will be used for which prompt, you can do it as follows:

```{r}
prompt_to_execute <- "Tell me about the economic situation in Algeria, summarize it in 3 bullet points, then translate it into German."

plan <- lead_agent$generate_plan(prompt_to_execute)
plan
```

Now, in order now to execute the workflow, we just need to call the `invoke` method which will behind the scene delegate the prompts to suitable Agents and retrieve back the final information: 

```{r}
response <- lead_agent$invoke("Tell me about the economic situation in Algeria, summarize it in 3 bullet points, then translate it into German.")
```

```{r}
response
```

If you want to inspect the multi-agents orchestration, you have access to the `agents_interaction` object: 

```{r}
lead_agent$agents_interaction
```


The above example is extremely simple, the usefulness of `mini007` would shine in more complex processes where a multi-agent sequential orchestration has a higher value added. 


## Visualizing agent plans with **visualize_plan()**

Sometimes, before running your workflow, it is helpful to view the orchestration as a visual diagram, showing the sequence of agents and which prompt each will receive. After generating a plan, you can call `visualize_plan()`:

This function displays the agents in workflow order as labeled boxes. Hovering a box reveals the delegated prompt. The visualization uses the `DiagrammeR` package. If no plan exists, it asks you to generate one first.

```{r, eval=FALSE}
lead_agent$visualize_plan()
```


## Broadcasting 

If you want to compare several `LLM` models, the `LeadAgent` provides a `broadcast` method that allows you to send a prompt to several different agents and get the result for each agent back in order to make a comparison and potentially choose the best agent/model for the defined prompt: 

Let's go through an example: 

```{r}
openai_4_1 <- ellmer::chat(
  name = "openai/gpt-4.1",
  credentials = retrieve_open_ai_credential, 
  echo = "none"
)

openai_4_1_agent <- Agent$new(
  name = "openai_4_1_agent", 
  instruction = "You are an AI assistant. Answer in 1 sentence max.", 
  llm_object = openai_4_1
)

openai_4_1_nano <- ellmer::chat(
  name = "openai/gpt-4.1-nano",
  credentials = retrieve_open_ai_credential, 
  echo = "none"
)

openai_4_1_nano_agent <- Agent$new(
  name = "openai_4_1_nano_agent", 
  instruction = "You are an AI assistant. Answer in 1 sentence max.", 
  llm_object = openai_4_1_nano
)
```

```{r}

lead_agent$clear_agents() # removing previous agents
lead_agent$register_agents(c(openai_4_1_agent, openai_4_1_nano_agent))
```


```{r}
lead_agent$broadcast(prompt = "If I were Algerian, which song would I like to sing when running under the rain? how about a flower?")
```

You can also access the history of the `broadcasting` using the `broadcast_history` attribute: 


```{r}
lead_agent$broadcast_history
```

## Human In The Loop (HITL)

When executing an LLM workflow that relies on many steps, you can set `Human In The Loop` (`HITL`) trigger that will check the model's response at a specific step. You can define a `HITL` trigger after defining a `LeadAgent` as follows:

```{r}
openai_llm_object <- ellmer::chat(
  name = "openai/gpt-4.1-mini",
  credentials = retrieve_open_ai_credential, 
  echo = "none"
)

lead_agent <- LeadAgent$new(
  name = "Leader", 
  llm_object = openai_llm_object
)

lead_agent$set_hitl(steps = 1)

lead_agent$hitl_steps
```
After setting the `HITL` to step 1, the workflow execution will pose and give the user 3 choices: 

1. Continue the execution of the workflow as it is;
2. Change manually the answer of the specified step and continue the execution of the workflow;
3. Stop the execution of the workflow (hard error);

Note that you can set a `HITL` at several steps, for example `lead_agent$set_hitl(steps = c(1, 2))` will set the `HITL` at step 1 and step 2. 


## Judge as a decision process 

Sometimes you want to send a prompt to several agents and pick the best answer. In order to choose the best prompt, you can also rely on the `Lead` Agent which will act a dudge and pick for you the best answer. You can use the `judge_and_choose_best_response` method as follows: 

```{r}
openai_4_1 <- ellmer::chat(
  name = "openai/gpt-4.1",
  credentials = retrieve_open_ai_credential, 
  echo = "none"
)

stylist_1 <- Agent$new(
  name = "stylist",
  instruction = "You are an AI assistant. Answer in 1 sentence max.",
  llm_object = openai_4_1
)

openai_4_1_nano <- ellmer::chat(
  name = "openai/gpt-4.1-nano",
  credentials = retrieve_open_ai_credential, 
  echo = "none"
)

stylist_2 <- Agent$new(
  name = "stylist2",
  instruction = "You are an AI assistant. Answer in 1 sentence max.",
  llm_object = openai_4_1_nano
)

openai_4_1_mini <- ellmer::chat(
  name = "openai/gpt-4.1-mini",
  credentials = retrieve_open_ai_credential, 
  echo = "none"
)

stylist_lead_agent <- LeadAgent$new(
  name = "Stylist Leader",
  llm_object = openai_4_1_mini
)

stylist_lead_agent$register_agents(c(stylist_1, stylist_2))

best_answer <- stylist_lead_agent$judge_and_choose_best_response(
  "what's the best way to wear a blue kalvin klein shirt in winter with a pink pair of trousers?"
)

best_answer
```

## Agents Dialog

The `agents_dialog` method facilitates an intelligent two-agent collaboration process designed to refine and optimize responses through iterative dialogue. 

It enables two registered agents to take alternating turns improving each other’s outputs until a high-quality final response is reached. The method supports a configurable maximum number of iterations (default: 5) and includes a self-stopping mechanism where agents can indicate agreement by beginning their message with “CONSENSUS:”, followed by the final answer. 

If no consensus is achieved within the iteration limit, the lead agent automatically synthesizes a concluding response based on the conversation. Throughout the exchange, every interaction is stored within the `self$dialog_history` object. Consider the following examples: 

```{r}
ceo1 <- Agent$new(
  name = "ceo1",
  instruction = paste0(
    "You are a CEO in  a dates company based in Ouergla, Algeria, ", 
    "You want to boost their exports to Germany. "
  ),
  llm_object = openai_4_1_mini
)

ceo2 <- Agent$new(
  name = "ceo2",
  instruction = paste0(
    "You are the CEO of a dates company based in Ouergla, Algeria. ", 
    "You are considering starting a marketing compaign to boost the exports to Germany. "
  ),
  llm_object = openai_4_1_mini
)

lead_agent <- LeadAgent$new(
  name = "Leader",
  llm_object = openai_4_1_mini
)

lead_agent$register_agents(c(ceo1, ceo2))

result <- lead_agent$agents_dialog(
  prompt = "Propose a plan in 1 sentence max about a marketing strategy that will boost the export of dates to Germany for the next 2 years",
  agent_1_id = ceo1$agent_id,
  agent_2_id = ceo2$agent_id,
  max_iterations = 3
)

# Access the final response
result$final_response

# View the dialog history
result$dialog_history
```


If the instructions of the Agents differ fundamentally, they won't be able to find a consensus and the `LeadAgent` will take over and provide a final response: 


```{r}
ceo1 <- Agent$new(
  name = "ceo1",
  instruction = paste0(
    "You are a CEO in  a dates company based in Ouergla, Algeria, ", 
    "You want to boost their exports to Germany. ", 
    "You don't care about the budget. You want to spend as much as possible. "
  ),
  llm_object = openai_4_1_mini
)

ceo2 <- Agent$new(
  name = "ceo2",
  instruction = paste0(
    "You are the CEO of a dates company based in Ouergla, Algeria. ", 
    "You are considering starting a marketing compaign to boost the exports to Germany. ",
    "For you the marketing budget is super important and you don't want to spend too much. "
  ),
  llm_object = openai_4_1_mini
)

lead_agent <- LeadAgent$new(
  name = "Leader",
  llm_object = openai_4_1_mini
)

lead_agent$register_agents(c(ceo1, ceo2))

result <- lead_agent$agents_dialog(
  prompt = "Propose a plan in 1 sentence max about a marketing strategy that will boost the export of dates to Germany for the next 2 years",
  agent_1_id = ceo1$agent_id,
  agent_2_id = ceo2$agent_id,
  max_iterations = 3
)

# Access the final response
result$final_response

# View the dialog history
result$dialog_history
```
