---
title: "LeadAgent"
format:
  html:
    toc: true
execute:
  echo: true
  warning: false
  error: false
---

```{r}
library(mini007)

retrieve_open_ai_credential <- function() {
  Sys.getenv("OPENAI_API_KEY")
}

openai_4_1_mini <- ellmer::chat(
  name = "openai/gpt-4.1-mini",
  credentials = retrieve_open_ai_credential,
  echo = "none"
)
```

# LeadAgent

## Creating a multi-agents orchestraction

We can create as many Agents as we want; the `LeadAgent` will dispatch the
instructions to the agents and provide the final answer back. Let's create three
Agents, a `researcher`, a `summarizer` and a `translator`:

```{r}
researcher <- Agent$new(
  name = "researcher",
  instruction = "You are a research assistant. Your job is to answer factual questions with detailed and accurate information. Do not answer with more than 2 lines",
  llm_object = openai_4_1_mini
)

summarizer <- Agent$new(
  name = "summarizer",
  instruction = "You are agent designed to summarise a give text into 3 distinct bullet points.",
  llm_object = openai_4_1_mini
)

translator <- Agent$new(
  name = "translator",
  instruction = "Your role is to translate a text from English to German",
  llm_object = openai_4_1_mini
)
```

Now, the most important part is to create a `LeadAgent`:

```{r}
lead_agent <- LeadAgent$new(
  name = "Leader",
  llm_object = openai_4_1_mini
)
```

Note that the `LeadAgent` cannot receive an `instruction` as it has already the
necessary instructions.

Next, we need to assign the Agents to `LeadAgent`, we do it as follows:

```{r}
lead_agent$register_agents(c(researcher, summarizer, translator))

lapply(lead_agent$agents, function(x) {x$name})
```

Before executing your prompt, you can ask the `LeadAgent` to generate a plan so
that you can see which `Agent` will be used for which prompt, you can do it as
follows:

```{r}
prompt_to_execute <- "Tell me about the economic situation in Algeria, summarize it in 3 bullet points, then translate it into German."

plan <- lead_agent$generate_plan(prompt_to_execute)
plan
```

Now, in order now to execute the workflow, we just need to call the `invoke`
method which will behind the scene delegate the prompts to suitable Agents and
retrieve back the final information:

```{r}
response <- lead_agent$invoke("Tell me about the economic situation in Algeria, summarize it in 3 bullet points, then translate it into German.")
response
```

If you want to inspect the multi-agents orchestration, you have access to the
`agents_interaction` object:

```{r}
lead_agent$agents_interaction
```

The above example is extremely simple, the usefulness of `mini007` would shine
in more complex processes where a multi-agent sequential orchestration has a
higher value added.

## Visualizing agent plans with `visualize_plan()`

Sometimes, before running your workflow, it is helpful to view the orchestration
as a visual diagram, showing the sequence of agents and which prompt each will
receive. After generating a plan, you can call `visualize_plan()`:

```{r, eval=FALSE}
lead_agent$visualize_plan()
```

## Broadcasting

If you want to compare several `LLM` models, the `LeadAgent` provides a
`broadcast` method that allows you to send a prompt to several different agents
and get the result for each agent back in order to make a comparison and
potentially choose the best agent/model for the defined prompt:

```{r}
openai_4_1 <- ellmer::chat(
  name = "openai/gpt-4.1",
  credentials = retrieve_open_ai_credential,
  echo = "none"
)

openai_4_1_agent <- Agent$new(
  name = "openai_4_1_agent",
  instruction = "You are an AI assistant. Answer in 1 sentence max.",
  llm_object = openai_4_1
)

openai_4_1_nano <- ellmer::chat(
  name = "openai/gpt-4.1-nano",
  credentials = retrieve_open_ai_credential,
  echo = "none"
)

openai_4_1_nano_agent <- Agent$new(
  name = "openai_4_1_nano_agent",
  instruction = "You are an AI assistant. Answer in 1 sentence max.",
  llm_object = openai_4_1_nano
)

lead_agent$clear_agents() # removing previous agents
lead_agent$register_agents(c(openai_4_1_agent, openai_4_1_nano_agent))

lead_agent$broadcast(prompt = "If I were Algerian, which song would I like to sing when running under the rain? how about a flower?")
```

You can also access the history of the `broadcasting` using the
`broadcast_history` attribute:

```{r}
lead_agent$broadcast_history
```

## Human In The Loop (HITL)

When executing an LLM workflow that relies on many steps, you can set `Human In
The Loop` (`HITL`) trigger that will check the model's response at a specific
step. You can define a `HITL` trigger after defining a `LeadAgent` as follows:

```{r}
openai_llm_object <- ellmer::chat(
  name = "openai/gpt-4.1-mini",
  credentials = retrieve_open_ai_credential,
  echo = "none"
)

lead_agent <- LeadAgent$new(
  name = "Leader",
  llm_object = openai_llm_object
)

lead_agent$set_hitl(steps = 1)

lead_agent$hitl_steps
```

After setting the `HITL` to step 1, the workflow execution will pose and give
the user 3 choices:

1. Continue the execution of the workflow as it is;
2. Change manually the answer of the specified step and continue the execution
   of the workflow;
3. Stop the execution of the workflow (hard error);

Note that you can set a `HITL` at several steps, for example
`lead_agent$set_hitl(steps = c(1, 2))` will set the `HITL` at step 1 and step 2.

