---
title: "Testing"
format:
  html:
    toc: true
execute:
  echo: true
  warning: false
  error: false
---

The non-deterministic behavior of LLM models makes `unit testing` impossible, nevertheless, there are ways to determine with a certain degree of a certainty the level of quality of a model. In this approach, we rely on the principle of `LLM as judge` to create `tests`. These tests can be evaluated at anytime using the `testthat` package.

Using the `agent$validate_response` function, it is possible to test the outputs produced by LLM agents. This is extremely useful if you want to test the responses given by a specific model. Note that the agent responsible for validating the responses can use any model supported by `ellmer`, meaning the `Agent` that generated the response and the `Agent` validating it don't have to share the same underlying model.  


## Example 

First we need to create an `Agent` responsible for testing our response:

```{r}
library(mini007)

retrieve_open_ai_credential <- function() {
  Sys.getenv("OPENAI_API_KEY")
}

openai_4_1_mini <- ellmer::chat(
  name = "openai/gpt-4.1-mini",
  credentials = retrieve_open_ai_credential,
  echo = "none"
)

tester <- Agent$new(
  name = "tester",
  instruction = "You are agent responsible of testing LLM responses",
  llm_object = openai_4_1_mini
)
```

Then, testing the response is relatively straightforward, not that in the following example, I've hard coded the `response` argument, however in practical application, it should come from an LLM agent.

```{r}
testthat::test_that("LLM provides good answers", {
  validation <- tester$validate_response(
    prompt = "What is the capital of Algeria",
    response = "The capital of algeria is Algiers", # This should come from an LLM Agent
    validation_criteria = "The capital of algeria should be Algiers",
    validation_score = 0.8
  )
  testthat::expect_true(validation$valid)
})
```
```{r, error=TRUE}
testthat::test_that("LLM provides good answers", {
  validation <- tester$validate_response(
    prompt = "What is the capital of Algeria",
    response = "The capital of algeria is Tokyo",
    validation_criteria = "The capital of algeria should be Algiers",
    validation_score = 0.8
  )
  testthat::expect_true(validation$valid)
})
```
