---
title: "Tools"
format:
  html:
    toc: true
execute:
  echo: true
  warning: false
  error: false
---

You can easily register one or several tools using the `register_tools` method. The tools are created using `ellmer`, consider the following example: 

```{r}
library(mini007)

openai_4_1 <- ellmer::chat(
  name = "openai/gpt-4.1",
  credentials = function() {Sys.getenv("OPENAI_API_KEY")},
  echo = "none"
)

weather_agent <- Agent$new(
  name = "weather_assistant",
  instruction = "You are a weather assistant.",
  llm_object = openai_4_1
)

weather_function_algiers <- function() {
  msg <- glue::glue(
    "35 degrees Celcius, it's sunny and there's no precipitation."
  )
  msg
}

get_weather_in_algiers <- ellmer::tool(
  fun = weather_function_algiers,
  name = "get_weather_in_algiers",
  description = "Provide the current weather in Algiers, Algeria."
)

weather_function_berlin <- function() {
  msg <- glue::glue(
    "10 degrees Celcius, it's cold"
  )
  msg
}

get_weather_in_berlin <- ellmer::tool(
  fun = weather_function_berlin,
  name = "get_weather_in_berlin",
  description = "Provide the current weather in Berlin, Germany"
)

weather_agent$register_tools(
  tools = list(
    get_weather_in_algiers, 
    get_weather_in_berlin
  )
)
```
One can list the available tools: 

```{r}
weather_agent$list_tools()
```

After registering the tools, the Agent will use them when needed: 

```{r}
weather_agent$invoke("How's the weather in Algiers?")
```

```{r}
weather_agent$invoke("How's the weather in Berlin?")
```

One can remove one or several tool using the `remove_tools` method or remove all agents at one using the `clear_tools` method: 

```{r}
weather_agent$clear_tools()
weather_agent$list_tools()
```
### Tool Generation

The `generate_and_register_tool` method allows you to create tools from simple natural language descriptions (for example, “create a tool that saves files to disk”) and automatically generates the complete R code needed to implement them. It produces a fully functional R function that encapsulates the tool’s logic, along with a complete `ellmer` tool definition that includes proper type specifications and clear parameter descriptions. 

```{r}
weather_agent$generate_and_register_tool(
  description = "create a tool that uses httr to call the open-meteo api https://open-meteo.com/en/docs to get the current weather about any city in the world"
)
```

```{r}
weather_agent$invoke(
  prompt = "what is the current weather in Tokyo?"
)
```
