---
title: "Workflow"
format:
  html:
    toc: true
execute:
  echo: true
  warning: false
  error: false
---

```{r}
library(mini007)

retrieve_open_ai_credential <- function() {
  Sys.getenv("OPENAI_API_KEY")
}

openai_4_1_mini <- ellmer::chat(
  name = "openai/gpt-4.1-mini",
  credentials = retrieve_open_ai_credential,
  echo = "none"
)

```

A `Workflow` is a predefined, sequential pipeline of processing units called **Stations**. Each Station receives the output of the previous one as its input, transforms it, and passes the result forward. **Routes** define which Station comes next and can be gated by optional condition functions that inspect the current output to decide the execution path.

Workflows complement `Agent` and `LeadAgent`:

- An `Agent` is LLM-driven and decides its own steps dynamically.
- A `LeadAgent` decomposes a prompt and delegates subtasks to agents automatically.
- A `Workflow` gives you **explicit control** over every step of the pipeline — you decide the order, the branching logic, and which handler runs where.

There are two main building blocks:

- `add_station(name, handler)` — registers a named processing unit. The handler can be an `Agent`, a `WorkflowAgent`, or a plain R `function`.
- `add_route(from, to)` — connects two Stations in sequence.

## A minimal linear pipeline

The simplest `Workflow` is a straight chain of Stations executed one after the other. Here we wire three specialised agents: a researcher that gathers facts, a writer that shapes them into a paragraph, and an editor that polishes the result.

```{r}
researcher <- Agent$new(
  name = "researcher",
  instruction = "You are a research assistant. Provide concise, accurate facts on the given topic in 3 max",
  llm_object  = openai_4_1_mini
)

writer <- Agent$new(
  name = "writer",
  instruction = "You are a skilled writer. Turn the research notes you receive into a single, engaging paragraph.",
  llm_object  = openai_4_1_mini
)

editor <- Agent$new(
  name = "editor",
  instruction = "You are a copy editor. Polish the paragraph you receive for grammar, clarity, and conciseness. Return only the final text.",
  llm_object  = openai_4_1_mini
)

translator <- Agent$new(
  name = "translator", 
  instruction = "You're a English-German Translater, translate text from English to German", 
  llm_object = openai_4_1_mini
)
```

```{r}
article_pipeline <- Workflow$new(
  name        = "article pipeline",
  description = "Research → Write → Edit -> Transalte", 
  use_cache = TRUE
)

article_pipeline$add_station(name = "research", handler = researcher, )
article_pipeline$add_station(name = "write", handler = writer)
article_pipeline$add_station(name = "edit", handler = editor)
article_pipeline$add_station(name = "translate", handler = translator)

article_pipeline$add_route(from = "research", to = "write")
article_pipeline$add_route(from = "write", to = "edit")
article_pipeline$add_route(from = "edit", to = "translate")

article_pipeline$set_entry(station_name = "research")
```

Calling `$run()` executes every Station in order. Each Station's output becomes the next Station's input automatically.

```{r}
result <- article_pipeline$run("The potential of Algeria when it comes to tourism in the Sahara")
result
```

The `run_history` attribute records every execution, including the full trace of inputs and outputs for each Station:

```{r}
article_pipeline$run_history
```

## Mixing agents with plain R functions

Stations do not have to be `Agent` objects. Any R function that accepts a single character argument and returns a character value is a valid handler. This makes it easy to add pre-processing or post-processing steps without an LLM call.

```{r}
# Plain R function stations, no LLM involved
normalise <- function(text) {
  trimws(gsub("\\s+", " ", text))
}

add_markdown_header <- function(text) {
  paste0("## Summary\n\n", text)
}

summariser <- Agent$new(
  name        = "summariser",
  instruction = "Summarise the text you receive into exactly three concise bullet points.",
  llm_object  = openai_4_1_mini
)
```

```{r}
format_pipeline <- Workflow$new("format and summarise")

format_pipeline$add_station(
  name = "normalise",
  handler = normalise, 
  description = "Collapse whitespace"
)
format_pipeline$add_station(
  name = "summarise", 
  handler = summariser, 
  description = "LLM bullet summary"
)
format_pipeline$add_station(
  name = "add_header", 
  handler = add_markdown_header, 
  description = "Prepend markdown header"
)

format_pipeline$add_route(from = "normalise",  to = "summarise")
format_pipeline$add_route(from = "summarise",  to = "add_header")
```

```{r}
messy_text <- "  The   2026    worldcup is    approachine.   Algeria's first game   is against     Argentina          "

format_pipeline$run(messy_text)
```

## Caching station results

When `use_cache = TRUE` (the default), every Station's output is stored keyed by its name and the input it received. A subsequent `$run()` call with the same input skips re-invoking any handler and returns the cached result immediately — saving both time and LLM costs when you are iterating on later Stations.

```{r}
cached_pipeline <- Workflow$new(
  name      = "cached-pipeline",
  use_cache = TRUE
)

drafter <- Agent$new(
  name        = "drafter",
  instruction = "Write a two-sentence draft on the topic provided.",
  llm_object  = openai_4_1_mini
)

refiner <- Agent$new(
  name        = "refiner",
  instruction = "Improve the draft you receive. Make it more vivid and precise.",
  llm_object  = openai_4_1_mini
)

cached_pipeline$add_station(name = "draft", handler = drafter)
cached_pipeline$add_station(name = "refine", handler = refiner)
cached_pipeline$add_route(from = "draft", to = "refine")
```

```{r}
# First run — both stations are executed
start <- Sys.time()
first_run <- cached_pipeline$run("Rare fish in the Algerian see")
first_run
end <- Sys.time()
print(end - start)
```

```{r}
# Second run with the same input — cache is served, no LLM calls are made
start <- Sys.time()
second_run <- cached_pipeline$run("Rare fish in the Algerian see")
second_run
end <- Sys.time()
print(end - start)
```

To force a fresh execution, call `$clear_cache()`:

```{r}
cached_pipeline$clear_cache()

# Fresh run after clearing the cache
start <- Sys.time()
fresh_run <- cached_pipeline$run("Rare fish in the Algerian see")
fresh_run
end <- Sys.time()
print(end - start)
```

## Conditional routing

Routes can carry a `condition`, a function that inspects the current Station's output and returns `TRUE` or `FALSE`. When a Station has multiple outgoing Routes, conditional ones are evaluated first (in the order they were added). The first condition that returns `TRUE` determines the next Station. If none match, the first unconditional Route is used as the default fallback.

This makes it possible to build branching pipelines where the path taken depends on what an earlier Station produces.

```{r}
classifier <- Agent$new(
  name        = "classifier",
  instruction = 'Classify the sentiment of the text. Reply with exactly one word: "positive" or "negative".',
  llm_object  = openai_4_1_mini
)

positive_handler <- Agent$new(
  name        = "positive_handler",
  instruction = "Write a warm, enthusiastic one-sentence reply to the following positive message.",
  llm_object  = openai_4_1_mini
)

negative_handler <- Agent$new(
  name        = "negative_handler",
  instruction = "Write an empathetic, solution-focused one-sentence reply to the following negative message.",
  llm_object  = openai_4_1_mini
)
```

```{r}
sentiment_router <- Workflow$new("sentiment router")

sentiment_router$add_station(name = "classify",  handler = classifier)
sentiment_router$add_station(name = "reply_pos", handler = positive_handler)
sentiment_router$add_station(name = "reply_neg", handler = negative_handler)

# Conditional routes — evaluated against the classifier's output
sentiment_router$add_route(
  from      = "classify",
  to        = "reply_pos",
  condition = function(out) grepl("positive", tolower(out))
)
sentiment_router$add_route(
  from      = "classify",
  to        = "reply_neg",
  condition = function(out) grepl("negative", tolower(out))
)

sentiment_router$set_entry(station_name = "classify")
```

```{r}
sentiment_router$run("I absolutely love this product, it changed my life!")
```

```{r}
sentiment_router$run("Very disappointed. The quality is nothing like advertised.")
```

A more structured example uses a plain function as the router so the branching logic is deterministic and free of LLM calls.

**Important pattern:** the router station must be a **passthrough** — it returns the original text unchanged so that downstream agents receive the real content. The routing classification belongs in the **route conditions**, which inspect the station output and decide where it goes next.

```{r}
brief_expander <- Agent$new(
  name        = "brief_expander",
  instruction = "The input is short. Expand it into a well-rounded paragraph of 2 sentences.",
  llm_object  = openai_4_1_mini
)

long_summariser <- Agent$new(
  name        = "long_summariser",
  instruction = "The input is long. Distill it into a single crisp sentence.",
  llm_object  = openai_4_1_mini
)

length_router <- Workflow$new("length router")

# The router station is a passthrough: it returns the text as-is.
# The length check lives in the route conditions below, not here.
length_router$add_station(name = "router", handler = function(text) text)
length_router$add_station(name = "short-to-long",  handler = brief_expander)
length_router$add_station(name = "long-to-short",   handler = long_summariser)

# Conditions receive the router's output (the original text) and decide the branch.
length_router$add_route(
  from      = "router",
  to        = "short-to-long",
  condition = function(x) nchar(x) < 80
)
length_router$add_route(
  from      = "router",
  to        = "long-to-short",
  condition = function(x) nchar(x) >= 80
)

length_router$set_entry(station_name = "router")
```

```{r}
# Short input — routed to the expander
length_router$run("Fennec Algeria football team")
```

```{r}
# Long input — routed to the summariser
long_input <- paste(
  "Bees are among the most important pollinators on the planet.",
  "They transfer pollen from one flower to another, enabling plants to reproduce.",
  "Without bees, many of the fruits and vegetables we eat would disappear.",
  "Colony collapse disorder has threatened bee populations worldwide,",
  "raising serious concerns about food security and ecosystem stability."
)
length_router$run(long_input)
```

## Wrapping a workflow as an agent

`$as_agent()` converts any `Workflow` into a `WorkflowAgent`. The result exposes the same `$invoke()` interface as a regular `Agent`, which means it can be:

- **registered with a `LeadAgent`** as one of its sub-agents, or
- **used as a Station handler** inside another `Workflow`.

This lets you compose complex pipelines from simpler ones without rewriting any logic.

### Using a WorkflowAgent inside a LeadAgent

```{r}
# Inner workflow: research + summarise
research_wf <- Workflow$new("research workflow")

gatherer <- Agent$new(
  name = "gatherer",
  instruction = "Gather three key facts about the topic provided. Be concise.",
  llm_object  = openai_4_1_mini
)

condonser <- Agent$new(
  name = "condenser",
  instruction = "Condense the facts you receive into a single sentence.",
  llm_object  = openai_4_1_mini
)

research_wf$add_station(name = "gather", handler = gatherer)
research_wf$add_station(name = "condense", handler = condonser)
research_wf$add_route(from = "gather", to = "condense")

research_wf$set_entry(station_name = "gather")
# Wrap as a WorkflowAgent
research_agent <- research_wf$as_agent(
  name = "research_agent",
  instruction = "An agent that gathers and condenses facts on any topic into one sentence."
)
```

```{r}
# Use the WorkflowAgent directly
research_agent$invoke("The city of Bejaia in Algeria")
```

```{r}
# Register with a LeadAgent alongside regular agents
translator <- Agent$new(
  name        = "translator",
  instruction = "Translate the text you receive from English into German",
  llm_object  = openai_4_1_mini
)

lead <- LeadAgent$new(
  name       = "Lead",
  llm_object = openai_4_1_mini
)

lead$register_agents(c(research_agent, translator))

lead$invoke("Tell me about the city of Bejaia in one sentence, then translate it into German")
```

### Nesting a WorkflowAgent inside another Workflow

```{r}
# Outer workflow: use the wrapped inner workflow as a Station
presentation_wf <- Workflow$new("Presentation workflow")

presentation_wf$add_station(
  name    = "research",
  handler = research_agent  # WorkflowAgent used as a station handler
)

presenter <- Agent$new(
    name = "presenter",
    instruction = "Turn the one-sentence summary you receive into a polished three-sentence report.",
    llm_object  = openai_4_1_mini
  )

presentation_wf$add_station(
  name = "present",
  handler = presenter
)

presentation_wf$add_route(from = "research", to = "present")
presentation_wf$set_entry(station_name = "research")

presentation_wf$run("Algeria and Coffee")
```

## Human In The Loop (HITL)

`$set_hitl(steps)` lets you pause the workflow at specific steps so a human
can inspect the Station's output before execution continues. Steps are
numbered from 1 in execution order — the same counter shown in the `run()`
console output.

When execution reaches a HITL step, it prints the Station name, the input it
received, and the output it produced, then offers three choices:

1. **Continue** — use the output as-is and move to the next Station.
2. **Edit** — type a replacement output; that value is forwarded to the next
   Station (and cached, if caching is on).
3. **Stop** — abort the workflow immediately.

HITL only fires on fresh executions; cache hits are skipped since the result
has already been reviewed in a prior run.

```{r}
hitl_pipeline <- Workflow$new("hitl pipeline", use_cache = FALSE)

drafter <- Agent$new(
  name = "drafter",
  instruction = "Write a short two-sentence paragraph on the topic provided.",
  llm_object = openai_4_1_mini
)

translator <- Agent$new(
  name = "translator",
  instruction = "Translate the text you receive from English into Spanish.",
  llm_object = openai_4_1_mini
)

hitl_pipeline$add_station(name = "draft", handler = drafter)
hitl_pipeline$add_station(name = "translate", handler = translator)

hitl_pipeline$add_route(from = "draft", to = "translate")

# Pause after step 1 (the drafter) so a human can review the draft
# before it is sent to the translator
hitl_pipeline$set_hitl(steps = 1)
```

```{r}
#| eval: false
hitl_pipeline$run("The discovery of penicillin")
```

You can set HITL at multiple steps at once:

```{r}
#| eval: false
hitl_pipeline$set_hitl(steps = c(1, 2), mode = "interactive")
```

### Non-blocking HITL: pause and resume (Not yet on CRAN)

The `HITL` above with `mode="interactive"` (the default) only works in an interactive console. Pass `mode = "pause"` to `$set_hitl()` and `$run()` returns immediately instead of
blocking, handing you back a `mini007_pending` object that describes exactly
where execution stopped. This is what lets HITL work from a Shiny app, a
Plumber API, or any batch job: persist the pending object, resume it whenever
the human actually responds, even in a different R session.

Let's route a teaser about Tassili n'Ajjer, the Algerian Sahara's
prehistoric rock-art plateau through a drafter and a translator, pausing
after the draft for review:

```{r}
sahara_pipeline <- Workflow$new("sahara pipeline", use_cache = FALSE)

drafter <- Agent$new(
  name = "sahara_drafter",
  instruction = "Write a punchy two-sentence teaser about the Saharan destination provided.",
  llm_object = openai_4_1_mini
)

translator <- Agent$new(
  name = "translator",
  instruction = "Translate the text you receive from English into German.",
  llm_object = openai_4_1_mini
)

sahara_pipeline$add_station(name = "draft", handler = drafter)
sahara_pipeline$add_station(name = "translate", handler = translator)
sahara_pipeline$add_route(from = "draft", to = "translate")

# mode = "pause" instead of the default "console"
sahara_pipeline$set_hitl(steps = 1, mode = "pause")
```

```{r}
#| eval: true
pending <- sahara_pipeline$run("Tassili n'Ajjer, Algeria")
pending
```

```
── Pending HITL request - step 1 ──
Request id: "b3f1c2a0-..."
Station/Agent: draft
ℹ Input:
Tassili n'Ajjer, Algeria
ℹ Proposed output:
Deep in Algeria's Sahara, Tassili n'Ajjer hides 15,000 years of rock art
carved into sandstone cathedrals — a UNESCO World Heritage plateau most
travellers never set foot on.
! Call `$resume("b3f1c2a0-...", action = "continue"|"edit"|"abort")` to proceed.
```

`is_pending(pending)` returns `TRUE` for this object (and `FALSE` for a
normal completed result). Once the human has weighed in, call `$resume()` with one of three actions:

```{r}
#| eval: false
# 1. Accept the draft as-is — the translator receives it next
sahara_pipeline$resume(pending$request_id, action = "continue")
```

```{r}
#| eval: false
# 2. Or edit it first. The edited text is what the translator receives
sahara_pipeline$resume(
  pending$request_id,
  action = "edit",
  value  = "Tassili n'Ajjer: 15,000 years of rock art etched into Saharan sandstone, in Algeria's far southeast."
)
```

```{r}
#| eval: false
# 3. Or abort altogether — raises an error, nothing downstream runs
sahara_pipeline$resume(pending$request_id, action = "abort")
```

If the workflow has more HITL steps ahead, `$resume()` returns another
`mini007_pending` object instead of the final result — so a caller typically
loops `while (is_pending(result))`, persisting `result$request_id` between
iterations, until a plain character string comes back.

Note that a `Workflow` wrapped as a `WorkflowAgent` (via `$as_agent()`) must
keep `mode = "console"` (the default): a `WorkflowAgent` is invoked
synchronously from inside a `Workflow` Station or a `LeadAgent` step, and
neither knows how to hand a paused run back to you — `$invoke()` raises an
error if the wrapped workflow pauses.

## Visualising a workflow

Call `$visualize()` on any `Workflow` to render an interactive directed graph. Stations appear as rounded blue boxes, conditional Routes as dashed arrows labelled "cond", and the entry Station is reached via a green **START** node.

```{r, eval=TRUE}
article_pipeline$visualize()
```

```{r, eval=TRUE}
sentiment_router$visualize()
```

The graph updates automatically to reflect whatever Stations and Routes are currently registered, making it a convenient tool for checking that the pipeline is wired correctly before running it.

---

## Per-station retry (Not yet on CRAN)

Real-world pipelines talk to external services — LLM providers, APIs, databases — that can fail transiently: rate-limit errors, network blips, brief provider outages. Rather than letting a single transient failure abort the entire run, you can attach a **retry logic** to any Station via two parameters of `$add_station()`:

| Parameter | Default | Meaning |
|-----------|---------|---------|
| `max_retries` | `0` | Extra attempts after the first failure |
| `retry_delay` | `1` | Seconds to sleep between attempts |

The station is tried up to `max_retries + 1` times in total. If it succeeds on any attempt, execution continues normally. If all attempts fail, the error is either forwarded to a [fallback handler](#fallback-handlers) (see below) or re-raised.

### Simulating a flaky service

```{r}
# Factory that returns a handler which fails the first `fail_times` calls
make_flaky_service <- function(fail_times = 2L) {
  attempts <- 0
  function(text) {
    attempts <<- attempts + 1L
    if (attempts <= fail_times) {
      stop(sprintf("HTTP 429: Rate limit (attempt %d of %d)", attempts, fail_times + 1L))
    }
    paste("Service result:", toupper(text))
  }
}
```

```{r}
#| message: false
retry_wf <- Workflow$new("retry pipeline", use_cache = FALSE)

retry_wf$add_station(
  name        = "call_service",
  handler     = make_flaky_service(fail_times = 2L),
  max_retries = 3,
  retry_delay = 0     # set to a positive number (e.g. 2) in production
)

retry_wf$run("summarise this document")
```

The station fails on attempts 1 and 2, then succeeds on attempt 3. The workflow returns the successful result transparently.

### Attaching retries to an Agent station

`max_retries` works identically for Agent stations — the LLM call is retried the same way.

```{r}
#| eval: false
fragile_pipeline <- Workflow$new("fragile pipeline")

fragile_pipeline$add_station(
  name        = "analyse",
  handler     = some_agent,      # any Agent or WorkflowAgent
  max_retries = 2,
  retry_delay = 3                # wait 3 s between retries
)
```

### Behaviour when all retries are exhausted

When every attempt fails and no [fallback](#fallback-handlers) is defined, the last error is re-raised and propagates up from `$run()`.

```{r}
always_fails_wf <- Workflow$new("always-fails pipeline", use_cache = FALSE)

always_fails_wf$add_station(
  name        = "broken_step",
  handler     = function(x) stop("Service permanently unavailable"),
  max_retries = 1,
  retry_delay = 0
)

tryCatch(
  always_fails_wf$run("input"),
  error = function(e) paste("Caught:", conditionMessage(e))
)
```

---

## Fallback handlers {#fallback-handlers} (not yet on CRAN)

A **fallback** is a `function(input, error)` attached to a specific Station. It is invoked only after all retry attempts for that Station have been exhausted, giving you a controlled degradation path instead of a hard failure.

```r
wf$add_station(
  name     = "step",
  handler  = primary_handler,
  fallback = function(input, error) {
    # `input`  — the string the station received
    # `error`  — the condition object from the last failed attempt
    "default output"
  }
)
```

The fallback's return value is used as the station's output and the pipeline continues normally from there.

### Simple graceful degradation

```{r}
safe_wf <- Workflow$new("safe pipeline", use_cache = FALSE)

safe_wf$add_station(
  name     = "fetch_summary",
  handler  = function(x) stop("Connection refused"),
  fallback = function(input, err) {
    paste("(Summary unavailable — service error:", conditionMessage(err), ")\nOriginal:", input)
  }
)

safe_wf$run("Analyse this quarterly report")
```

### Fallback fires only after all retries

The fallback is a last resort. With `max_retries > 0` the primary handler is tried repeatedly; the fallback only runs if every attempt fails.

```{r}
#| message: false
attempts <- 0L

resilient_wf <- Workflow$new("resilient pipeline", use_cache = FALSE)

resilient_wf$add_station(
  name        = "external_call",
  handler     = function(x) {
    attempts <<- attempts + 1L
    stop(sprintf("Attempt %d failed", attempts))
  },
  max_retries = 2,
  retry_delay = 0,
  fallback    = function(input, err) {
    paste0("[Fallback after ", attempts, " attempt(s)] ", input)
  }
)

resilient_wf$run("Process this document")
```

### Routing continues after a fallback

The fallback's output flows into the next Station via the normal routing mechanism. The rest of the pipeline is unaware that a fallback fired.

```{r}
routed_fallback_wf <- Workflow$new("routed fallback", use_cache = FALSE)

routed_fallback_wf$add_station(
  name     = "fetch",
  handler  = function(x) stop("API down"),
  fallback = function(input, err) paste("CACHED VERSION OF:", input)
)

routed_fallback_wf$add_station(
  "format",
  function(x) paste0("**", trimws(x), "**")
)

routed_fallback_wf$add_route("fetch", "format")

routed_fallback_wf$run("monthly sales figures")
```

### Using a backup LLM as the fallback

A common production pattern is to call a premium model as the primary handler and fall back to a cheaper or local model when the primary is unavailable.

```{r}
#| eval: false
primary_agent  <- Agent$new("primary",  "You are a senior analyst.", premium_llm)
fallback_agent <- Agent$new("fallback", "You are a concise analyst.", cheap_llm)

pipeline <- Workflow$new("model-with-backup")

pipeline$add_station(
  name     = "analyse",
  handler  = primary_agent,
  max_retries = 1,
  retry_delay = 2,
  fallback = function(input, err) {
    fallback_agent$invoke(input)   # delegate to the backup model
  }
)
```

---

## Parallel Stations (not yet on CRAN)

Multiple independent Stations can run concurrently using `$set_daemons()` and `$add_parallel_group()`. 

### Multi-perspective parallel analysis

```{r}

# Create three specialized agents for different perspectives
technical_agent <- Agent$new(
  name = "technical_analyst",
  instruction = paste(
    "You are a technical expert in AI and software engineering.",
    "Analyze topics from a technical/engineering perspective.",
    "Focus on: implementation details, technology stack, technical challenges.",
    "Keep responses to 2-3 sentences."
  ),
  llm_object = openai_4_1_mini
)

business_agent <- Agent$new(
  name = "business_strategist",
  instruction = paste(
    "You are a business strategist and market analyst.",
    "Analyze topics from a business/market perspective.",
    "Focus on: market opportunities, ROI, competitive advantage, business models.",
    "Keep responses to 2-3 sentences."
  ),
  llm_object = openai_4_1_mini
)

ux_agent <- Agent$new(
  name = "ux_specialist",
  instruction = paste(
    "You are a UX and product design specialist.",
    "Analyze topics from a user experience perspective.",
    "Focus on: user needs, design considerations, accessibility, user satisfaction.",
    "Keep responses to 2-3 sentences."
  ),
  llm_object = openai_4_1_mini
)


# Create workflow with parallel execution
wf <- Workflow$new(
  name = "parallel_multi_agent_analysis",
  description = "Multi-perspective analysis using parallel Agent invocations"
)

# Set up 3 parallel workers for concurrent API calls
wf$set_daemons(3)

# Add input station (passthrough)
wf$add_station(
  "input",
  handler = function(x) x,
  description = "Input topic"
)

# Add three parallel Agent stations
wf$add_station(
  "technical_analysis",
  handler = technical_agent,
  description = "Technical perspective"
)

wf$add_station(
  "business_analysis",
  handler = business_agent,
  description = "Business perspective"
)

wf$add_station(
  "ux_analysis",
  handler = ux_agent,
  description = "UX perspective"
)

# Add synthesis station
wf$add_station(
  "synthesis",
  handler = function(x) {
    paste0(
      "═══════════════════════════════════════════\n",
      "MULTI-PERSPECTIVE ANALYSIS REPORT\n",
      "═══════════════════════════════════════════\n\n",
      x
    )
  },
  description = "Synthesis"
)

# Define parallel group: input -> [technical, business, ux] -> synthesis
wf$add_parallel_group(
  from = "input",
  stations = c("technical_analysis", "business_analysis", "ux_analysis"),
  to = "synthesis",
  merge_fn = function(results) {
    paste0(
      "📱 TECHNICAL PERSPECTIVE:\n",
      results$technical_analysis, "\n\n",
      "💼 BUSINESS PERSPECTIVE:\n",
      results$business_analysis, "\n\n",
      "👤 UX PERSPECTIVE:\n",
      results$ux_analysis
    )
  }
)

wf$visualize()
```



```{r}
# Execute workflow with parallel execution
topic <- "Artificial Intelligence in Healthcare"

result <- wf$run(topic)

# Clean up any existing daemons from previous runs
mirai::daemons(0)

result
```



### How it works

1. **`$set_daemons(n)`** Creates `n` parallel worker processes.

2. **`$add_parallel_group(from, stations, to, merge_fn)`** Defines which stations run in parallel:
   - `from`: predecessor station (all parallel stations receive its output)
   - `stations`: vector of station names to run concurrently
   - `to`: successor station (receives merged results)
   - `merge_fn`: function to combine results from all parallel stations

3. **Execution**  All parallel stations run concurrently in separate daemons

4. **Visualization** `$visualize()` shows parallel routes in **red with ∥ symbol** to distinguish from sequential routes

